---
output: github_document
always_allow_html: true
---

This is an introductory article that teaches users how to use the
zoomerjoin.extended R package to generate embeddings and perform a join on
them.

## Installation

## Zoomerjoin Development Version

Please note that `zoomerjoin.extended` at the moment depends on the
*development* version of the `zoomerjoin` package.  To install this package,
please use the following code:

```{r, eval = FALSE}
install.packages('zoomerjoin', repos = c('https://beniaminogreen.r-universe.dev', 'https://cloud.r-project.org'))
```

In the future, this dependency will be relaxed, and you will be able to install
all of the package dependencies that you need from CRAN.

```{r setup}
library(zoomerjoin.extended)
```

## Installing the ONNXRuntime

`zoomerjoin.extended` also uses Microsoft's [ONNX
Runtime](https://onnxruntime.ai/) to provide speedy inference for the neural
joins.  Please use the ONNX Runtime [getting started
guide](https://onnxruntime.ai/getting-started) to install the correct version
of the runtime. At the moment, only CUDA and CPU backends are supported.

Once this is setup, you may need to point the package to the location the ONNX
Runtime is installed. You can do this using the `ORT_DYLIB_PATH` environment
variable as follows:

```{r, eval=F}
Sys.setenv("ORT_DYLIB_PATH"="/path/to/onnx_runtime/")
```

```{r, echo=FALSE, include=FALSE, eval=TRUE}
Sys.setenv("ORT_DYLIB_PATH"="/nix/store/rrp6kppq6jlqj765bqkab2lr6mizhix0-onnxruntime-1.18.1/lib/libonnxruntime.so")
```

# Basic Tutorial

In this tutorial, I will discuss how to generate embeddings using the [popular all-MiniLM-L6-v2 model on
HuggingFace](https://huggingface.co/sentence-transformers/all-MiniLM-L6-v2).
The process will be substantially similar for other models trained for the
'sentence-similarity' task on the hugging face hub.

## Downloading Model Files

First, you must download the model files from the 'files and versions' pane for
the model on huggingface. To run the neural joins, you need the tokenizer files
(`tokenizer.json`, and `tokenizer_config.json`), and the model file in an ONNX
format. This huggingface repository has several versions of the model, and we
will pick the one called `model_O3.onnx`. This version of the model has some
optimizations to make it run faster, but can still be run on the CPU.

Download all these files, and put them into a diretory (in this case, we are
using `_model/`).

```
_model/
├── model_O3.onnx
├── tokenizer_config.json
└── tokenizer.json
```

Once you have the model files set up correctly, you can load the model into
memory, and start generating emeddings. You can do load the model the
`new_inference_session` function, which takes a path to the directory holding
the model and tokenizer, and also allows you specify which files hold the model
and tokenizer if they differ from the defaults ('model.onnx' and
'tokenizer.json', respectively).

```{r}
model <- new_inference_session("_model/", model = "model_O3.onnx")
model
```

Now you are ready to get going! You can generate embeddings for strings as follows:

```{r}
embeddings <- generate_embeddings(model, c("the quick brown fox", "jumped over the lazy dog"))
# only print first 5 dimensions
embeddings[, 1:5]
```

Or join datasets based on embeddings of a variables using a familiar, dplyr-style syntax.

```{r}
a <- data.frame(
               name_1 =
                   c("JP Morgan Chase",
                     "he was sad",
                     "he is happy",
                     "Bank of America",
                     "the lazy dog was jumped over by the quick brown fox"
                   )
               )
b <- data.frame(
                name_2 =
                    c("JP Morgan",
                      "the man was morose",
                      "the man was smiling",
                      "Bank of America",
                      "the quick brown fox vaulted over the sleeping dog")
               )
neural_inner_join(a,b, model, by = c('name_1' = 'name_2'), radius = .5)
```

